---
title: "p8105_hw6_ah4167"
author: "Aiying Huang"
date: "2023-11-19"
output: github_document
---
```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem1

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide_df <- read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown"))|>
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )
  ) |>
  mutate(
    is_solved = ifelse(resolution == "solved", 1, 0)
  )|>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) &
           (victim_race %in% c("White","Black")))|>
  mutate(
    victim_age=as.numeric(victim_age)
  )
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_data=homicide_df|>
  filter(city_state=="Baltimore, MD")|>
  select(is_solved,victim_age,victim_sex,victim_race)
model = glm(is_solved ~ victim_age + victim_sex + victim_race, data = baltimore_data, family = binomial)
model|>
  broom::tidy()|>
  filter(term == "victim_sexMale")
```
```{r}
# Calculate the confidence interval for male vs. female
estimate = model|>
  broom::tidy()|>
  filter(term == "victim_sexMale")|>
  pull(estimate)
std_error = model|>
  broom::tidy()|>
  filter(term == "victim_sexMale")|>
  pull(std.error)

# Set the confidence level (e.g., 95%)
confidence_level <- 0.95

# Calculate the critical value for the confidence interval
z <- qnorm((1 + confidence_level) / 2)

# Calculate the margin of error
margin_error <- z * std_error

# Calculate the lower and upper bounds of the confidence interval
lower_bound <- exp(estimate - margin_error)
upper_bound <- exp(estimate + margin_error)

# Print the estimate and its confidence interval
cat("Estimate:", exp(estimate), "\n")
cat("95% Confidence Interval: (", lower_bound, ", ", upper_bound, ")\n")

```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
# Load the necessary libraries
library(dplyr)
library(purrr)
library(broom)

# Create a list column for each city
city_list = homicide_df |>
  group_by(city_state) |>
  nest()

# Define a function to fit logistic regression models and extract ORs and CIs
fit_logistic_regression <- function(data) {
  model = glm(is_solved ~ victim_age + victim_sex + victim_race, data = data, family = binomial)
  model_tidy = tidy(model)
  male_vs_female = model_tidy |>
    filter(term == "victim_sexMale")
  estimate = male_vs_female$estimate
  std_error = male_vs_female$std.error
  z = qnorm((1 + 0.95) / 2)  # 95% confidence interval
  margin_error = z * std_error
  lower_bound = exp(estimate - margin_error)
  upper_bound = exp(estimate + margin_error)
  tibble(
    city_state = unique(data$city_state),
    estimate = exp(estimate),
    ci_lower = lower_bound,
    ci_upper = upper_bound
  )
}

# Use purrr::map to fit models and extract adjusted ORs and CIs for each city
results = city_list |>
  mutate(city_results = map(data, fit_logistic_regression))

# Unnest the results to create a tidy dataframe
tidy_results = results |>
  unnest(city_results)

# View the estimated ORs and CIs for each city
head(tidy_results)

```


Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
# Load the necessary libraries
library(ggplot2)

# Assuming you have a dataframe "tidy_results" with estimated ORs and CIs for each city
# Organize the cities by estimated OR in descending order
tidy_results |>
  arrange(desc(estimate))|>
  ggplot(aes(x = reorder(city_state, estimate), y = estimate, ymin = ci_lower, ymax = ci_upper))+
  geom_point() +
  geom_errorbar() +
  labs(
    x = "City State",
    y = "Adjusted Estimated Odds Ratio (OR)",
    title = "Adjusted Estimated Odds Ratios and Confidence Intervals by City"
  ) +
  coord_flip()  # Rotate x-axis labels for better readability



```




